WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2018-12-16 paper/ms.tex
update author list
merged JNW comments

\NewPageAfterKeywords

@@

@@

@@

\affiliation{ Harvard-Smithsonian Center for Astrophysics, 60 Garden

@@

\begin{abstract}
  By combining data from the Transiting Exoplanet Survey Satellite
  (TESS) with data from previous studies, we show that the transit
  times of WASP-4b are incompatible with a constant orbital period.
  The transits observed by TESS occurred $81.5 \pm 11.8$~seconds
  earlier than one would expect, based on the previous data.  The
  orbital period appears to be shrinking by $\dot{P}=-12.6 \pm 1.2$
  milliseconds per year.   A simple clock offset is unlikely, because
  TESS observations of other hot Jupiters (WASP-6b, 18b, and 46b) are
  compatible with a constant period and rule out a 81.5-second offset
  at the 6.3$\sigma$ level.  The timing variations seem to be
  astrophysical, and could be caused by apsidal precession, or perhaps
  by tidal decay.  The latter seems implausible on theoretical
  grounds, as it would require a stellar quality factor $Q_\star'
  \approx 3\times10^4$.  Apsidal precession would imply an
  eccentricity $e\approx1.8\times 10^{-3}$, and the possibility of
  measuring the planet's Love number.  A Doppler shift from WASP-4's
  acceleration towards us could account for at most one quarter of the
  observed period change (at $2\sigma$).  Further observations are
  needed to confirm and understand the timing variation.
\end{abstract}

@@

Long-term monitoring of hot Jupiter transit and occultation times
should eventually reveal two distinct processes.  First, whenever the
summed spin and orbital angular momenta in a hot Jupiter system fall
below a critical value, the system becomes unstable to tidal decay
\citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
Jupiters satisfy this condition, so their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  However the rate of decay
is uncertain, as it depends on how friction dissipates energy carried by
gravitational tides (\citealt{ogilvie_tidal_2014} gives a review).  If
we could directly measure orbital decay rates, it would inform our
understanding of tidal dissipation.  This would be important because the
dissipation rates determine the fate of the system.  During its
inspiral, if the planet fills its Roche lobe below the stellar
surface, the drag force from the stellar envelope accelerates the
inspiral, causing a `direct-impact' merger
\citep{metzger_optical_2012,macleod_planetary_2018} Alternatively, if
the planet fills its Roche lobe above the stellar surface, then
it could either be disrupted into an accretion disk or it could slowly
transfer mass to the star
\citep{metzger_optical_2012,valsecchi_tidally-driven_2015,jackson_tidal_2016}.
Further afield, the outcome of the eventual interaction between the
Earth and red giant Sun also hinges on the efficiency of tidal
dissipation \citep{rasio_tidal_1996}.

The second process that long-term timing studies should reveal is
apsidal precession, {\it i.e.}, rotation of the orbital ellipse within
the orbital plane. Apsidal precession has been observed extensively in
eclipsing binaries \citep[{\it e.g.},][]{
  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
In exoplanetary systems, precessing orbits have been detected in a
handful of cases, including circumbinary systems (Kepler-413
\citealt{kostov_kepler-413b_2014}; Kepler-453
\citealt{welsh_kepler_2015}, planets orbiting rapidly rotating stars
(Kepler-13
\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
PTFO 8-8695 \citealt{barnes_measurement_2013}), and systems with
multiple interacting giant planets (Kepler-108
\citealt{mills_kepler-108_2017}; Kepler-693
\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
observed in hot Jupiter systems, perhaps in part because their orbits
are nearly circular.  However, there are many mechanisms that could
pump hot Jupiter orbits away from circularity (see
\S~\ref{sec:apsidal_precession}, and
\citealt{bailey_understanding_2019}), which would then lead to
observable timing variations.  For hot Jupiters, the apsidal
precession is usually dominated by the quadrupolar distortion of the
planet by the star's tidal force \citep{ragozzine_probing_2009}.  
If apsidal precession were observed in a hot Jupiter system, the
precession rate could thus be used to compute the planet's Love
number, giving a third parameter along with the mass and radius with
which to describe the planet's interior structure \citep[{\it
e.g.},][who performed a similar procedure for
HAT-P-13b]{batygin_determination_2009}.  Measuring small
eccentricities and understanding their origin might also help us
understand the dynamical histories and inflated radii of hot Jupiters.
\citep[{\it
e.g.},][respectively]{dawson_origins_2018,ibgui_tidal_2010}

While numerous indirect studies have called attention to
population-level effects of tidal decay
\citep{jackson_observational_2009,hansen_calibration_2010,penev_constraining_2012,husnoo_observational_2012,matsakos_origin_2016,cameron_hierarchical_2018,penev_empirical_2018},
the most convincing direct evidence to date for either orbital decay
or apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} showed that the transit times of
WASP-12b did not follow a linear ephemeris at $5\sigma$ confidence,
and measured $\dot{P} = -26 \pm 4\,{\rm ms}\,{\rm yr}^{-1}$.  Using
new timing data, \citet{patra_2017} found $\dot{P} = -29 \pm 3\,{\rm
ms}\,{\rm yr}^{-1}$.  If the period change were caused entirely by
tidal decay, the implied modified stellar quality factor would be
$Q_\star' \approx 2\times10^5$.  Standard equilibrium tides do not
predict this level of dissipation
\citep{penev_tidal_2011,ogilvie_tidal_2014}.  The needed level of
friction might come from dynamical tides, or obliquity tides
\citep{weinberg_tidal_2017,millholland_obliquity_2018}.  The former is
more likely if WASP-12 is a subgiant, but the observational
constraints favor a main-sequence model
\citep{bailey_understanding_2019}.  More data are needed to understand
the WASP-12b timing variations.

This work focuses on the hot Jupiter WASP-4b, which orbits its host
star every 1.34 days, at a distance of $a/R_\star = 5.46$
\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  WASP-4b is a plausible
candidate for orbital decay, and has been
observed often over the past decade.  The host is a G7V dwarf, and is on the
main sequence (\S~\ref{sec:system_parameters}).  The orbit is
consistent with being circular (but see
\S~\ref{sec:apsidal_precession}), and having sky-projected obliquity
of zero
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.

We measure new transit times for WASP-4b using TESS lightcurves
(\S~\ref{sec:observations},~\ref{sec:measurement}), and calculate the
system's parameters empirically (\S~\ref{sec:system_parameters}).  We
merge the new times with previous observations and compare three
models (\S~\ref{sec:timing}): a constant period, a decaying period,
and a slightly eccentric orbit undergoing precession.  We rule out a
constant period, but cannot firmly distinguish between the latter two
possibilities.  Each would have interesting implications
(\S~\ref{sec:implications}).  We rule out the possibility that a
systematic time system offset causes the observed variation
(Appendix~\ref{sec:verify_tess}), and conclude by advocating for
further monitoring of the system (\S~\ref{sec:future}).

\section{New transits and system parameters}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Phase-folded transit of WASP-4b.} Black points are TESS
        flux measurements; yellow points are binned measurements.
        Using the fit to the phase-folded transit (blue line), we
        measure mid-transit times by fitting the individual transits
        (Figure~\ref{fig:lightcurves}).
        \label{fig:phasefold}
    }
\end{figure}

\begin{figure}[t]
    \begin{center}
        \includegraphics[width=0.48\textwidth]{f2.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Individual transits of WASP-4b.} On the left, black
        points are TESS flux measurements, with a vertical offset
        applied. Blue curves are best-fit models. Rounded midtimes are
        printed in BTJD next to the appropriate transits.  The
        residuals are shown on the right.
        \label{fig:lightcurves}
    }
\end{figure}

\subsection{Observations}
\label{sec:observations}

WASP-4 was observed in TESS camera 2 from August 23, 2018 to September
20, during the spacecraft's second sector of science operations.  The
star is designated by the TESS Input Catalog as TIC 402026209
\citep{stassun_TIC_2018}.

The TESS spacecraft recorded images of WASP-4 in an $11\times11$ pixel
array every 2 minutes.   The data were downlinked via the Deep Space
Network. The timestamps from the spacecraft clock were then converted
by the Payload Operations Center to TESS Julian Day in the {\it
Temps Dynamique Barycentrique} (TDB) reference.  The images were then
reduced to lightcurves by the Science Processing Operations Center (SPOC)
at NASA Ames~\citep{jenkins_tess_2016}.  During this processing, SPOC used
the known
spacecraft trajectory to compute the barycentric time corrections on a
target by target basis, and reported timestamps in ``BTJD'', where
${\rm BTJD} = {\rm BJD} - 2457000$, BJD is the barycentric julian
date, and all times are in the TDB reference (\citealt{urban_explanatory_2012} 
describe the relevant time systems).
The lightcurves were then vetted and released by the MIT TESS Science
Office to the Mikulski Archive for Space
Telescopes~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC) light
curves from MAST.  The steps used to find an optimal aperture (for
WASP-4, a $3\times 3$ pixel array) and the PDC-MAP algorithm are
described by \citet{smith_kepler_apertures_2017} and
\citet{smith_kepler_PDC_2017} respectively.

We processed the lightcurve as follows.  First, we removed all
points with non-zero quality flags.  This removed data contaminated by
coarse spacecraft pointing, cosmic rays, and other annoyances
\citep{tess_data_product_description_2018}.  We then filtered out
known bad observing windows.  The data closest to spacecraft perigee
typically show ramp-like systematics, so we clipped out the first and
last hour of both orbits.  Next, we focused on the momentum dumps. As
described in the data release
notes\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}},
during the first two TESS sectors, every 2.5~days the spacecraft's
reaction wheels were reset to maintain pointing stability.  These
events were assigned quality flags corresponding to ``Reaction Wheel
Desaturation Event'' and ``Manual Exclude''.  For WASP-4, these flags
were simultaneously set for 54 distinct cadences, and there were 10
momentum dumps, averaging about 10 minutes of flagged data per dump.
Simply out of caution, we clipped out an additional 10 minutes before
and after every momentum dump.  Before applying any filters, we had
19737 data points. After applying all the filters, we were left with
18165 measurements, or 92\% of the original data.

After ``cleaning'' the data, we normalized the flux measurements by
dividing out the median flux.  We converted the timestamps from BTJD
to BJD by adding the appropriate 2457000 day offset
\citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018} We did not
``flatten'' the lightcurves, as is often done with {\it
e.g.}, a spline or gaussian process regression.  Such procedures are
asymmetric about the transit minimum and can skew the midtime
measurement without being captured by the measurement uncertainties.

@@

\begin{figure}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.48\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.6cm}
    \caption{
        {\bf TESS saw WASP-4b transit earlier than expected.} Both
        plots show observed transit times minus the pre-TESS ephemeris
        as a function of transit number.  Blue curves are the $\pm
        1\sigma$ credible interval around the constant-period
        prediction.  {\it Top:} The full timing dataset spans 10
        years.  Less precise points are more transparent.  {\it
        Bottom:} The binned TESS observation (red band) arrived $81.5
        \pm 11.8\ {\rm seconds}$ before it was expected.
        \label{fig:arrived_early}
    }
\end{figure}

Using the cleaned PDC lightcurve, we ran Box Least Squares (BLS) to
estimate the orbital period, transit duration and a reference epoch
\citep{kovacs_box-fitting_2002}.  From these parameters, we isolated
each transit to within $\pm$4 transit durations of its midtime.  We
then fit the phase-folded transit, using the analytic formulae
calculated by \citet{mandel_analytic_2002} and implemented by
\citet{kreidberg_batman_2015}.  Specifically, we fit for a reference
epoch, the planet to star radius ratio $R_{\rm p}/R_\star$, the
semi-major axis to stellar radius ratio $a/R_\star$, the inclination
$i$, two quadratic limb-darkening coefficients $(u_{\rm linear},
u_{\rm quad})$, and the orbital period $P$.  We set the eccentricity
to zero.

Having defined the free and fixed parameters, we sampled over the
posterior using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
are in reasonable agreement with the parameters reported by {\it
e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.
Figure~\ref{fig:phasefold} shows the phase-folded lightcurve.

To measure the transit midtimes, we then fit  each individual transit
for a model transit, plus a local linear trend.  We allowed four free 
parameters per transit: the time of
mid-transit $t_{\rm tra}$, the planet-to-star radius ratio, and the
slope and intercept of the line.  We fixed the remaining
parameters%~---~$a/R_\star$, $i$, $P$, and the limb-darkening
%coefficients~---~
to the best-fit values determined from the
phase-folded TESS lightcurve.

After completing an initial fit for each transit, we set the
photometric error bars to equal the standard deviation of the
out-of-transit points of the residual lightcurve.  We then repeated
the fitting procedure, using the empirically determined photometric
errors.

To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 9.2$, with 16 degrees of
freedom.  Visually inspecting residuals showed that the error variance
had been overestimated, so we multiplied the measured TESS errors by a
factor $f=0.76$, forcing a reduced $\chi^2$ of unity.  This lowered
the mean uncertainty of the transit midtimes from $29.8$ to $22.6$
seconds.

Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
best-fit models, and the residuals.  Table~2 reports the mid-transit
times and their uncertainties.
Binning the residuals to 1-hour windows and taking the median absolute
deviation, the lightcurves have an RMS of $586\,{\rm ppm}\,{\rm
hr}^{1/2}$.  The predicted\footnote{\url{github.com/lgbouma/tnm}}
photon-counting noise is $410\,{\rm ppm}\,{\rm hr}^{1/2}$.  For a
$T\approx11.8$ star with a 9-pixel aperture, read noise is expected to
contribute an additional $202\,{\rm ppm}\,{\rm hr}^{1/2}$ in
quadrature.  The predicted zodiacal background contribution is
$673\,{\rm ppm}\,{\rm hr}^{1/2}$, based on the
\citet{winn_photonflux_2013} model, which \citet{Sullivan_2015} used
in planet detection simulations.  This zodiacal noise prediction seems
to have been an overestimate, given the observed RMS of WASP-4.  A
broader assessment of TESS's photometric performance is beyond
the scope of this work.

@@

\tablecomments{
  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
  Orbital periods are in Table~4. The limb darkening parameters were
  allowed to float around the \citet{claret_limb_2017} prediction, but
  were unconstrained.
  (B) \citet{triaud_spin-orbit_2010}.
  (C) From HARPS spectra \citep{doyle_accurate_2013}.
  (D) \citet{stassun_accurate_2017}.
  (E) From functions of stellar, transit or RV parameters as discussed in text.
  (F) {\it Gaia} DR2.
  (G) \citet{stassun_TIC_2018}.
}

@@

\citet{doyle_accurate_2013} found accurate spectroscopic parameters
for WASP-4 using high S/N HARPS spectra.  Using their (T$_{\rm eff}$,
$\log g$, $[{\rm Fe/H}]$) and the available {\it Gaia} parallaxes, we
calculate the remaining system parameters empirically.

We compute the bolometric flux using broadband magnitudes from
available all-sky catalogs, as follows: $G$ from {\it Gaia\/} DR2,
$B_T V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from {\it
2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
wavelength range 0.4--22~$\mu$m.  Combining the implied angular radius
with the {\it Gaia} DR2 parallax, and adjusting for the parallax
offset \citep{stassun_evidence_2018}, we obtain the stellar radius.
For the stellar mass, we started with the $a/R_\star$ value found from
the phase-folded TESS lightcurve.  The implied stellar density follows
\citep{seager_unique_2003}, as do the
semi-major axis and stellar mass.  We get the planet radius using our
$R_{\rm p}/R_\star$ measurement.  Finally, for the planet mass we use the
inclination angle from the transit, and the RV semi-amplitude from 
\citet{triaud_spin-orbit_2010}.

We list the resulting parameters in Table~1, and adopt them for the
remaining analysis.  Our parameters broadly agree with those of previous
authors
\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
% The sole exception is that our measured $R_{\rm p}/R_\star$ value is
% slightly smaller than most ground-based measurements (see Tables 2 and
% 3 of \citealt{hoyer_tramos_2013}).  However, looking at the I-band
% measurements of \citet{hoyer_tramos_2013} and taking the mean + std
% deviation, it's a ~1 sigma effect. So I won't mention it!
Comparing our empirical result with the Yonsei-Yale evolutionary track
predicted by the spectroscopic parameters, we see that WASP-4 has an
age $\approx 7\,{\rm Gyr}$, and is securely in the main-sequence stage
of evolution.

@@

@@

\begin{figure*}[t]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.9\textwidth]{f4.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Timing residuals and best-fit models for WASP-4b.}
        Points are observed transit times minus a linear fit to all
        observed times.  Less precise times are more transparent.  The
        best constant period model (gray line) is a poor fit to the
        observations.  The best-fit tidal decay model (blue) and
        apsidal precession model (orange) can fit the data.  The
        binned TESS point (red square) is for display purposes only;
        the individual TESS times are used to fit the models.
        \label{fig:times}
    }
\end{figure*}

>>>>

%\NewPageAfterKeywords

@@

\author[0000-0003-3438-843X]{C. Baxter}

@@

\author[0000-0002-0628-0088]{W. Bhatti}

@@

@@

\begin{abstract}
    The Transiting Exoplanet Survey Satellite (TESS) recently observed
    18 transits of the hot Jupiter WASP-4b.  The sequence of transits
    occurred $81.5 \pm 11.8$~seconds earlier than had been predicted,
    based on earlier data stretching back to 2007.  This is unlikely
    to be the result of a clock error, because TESS observations of
    other hot Jupiters (WASP-6b, 18b, and 46b) are compatible with a
    constant period, ruling out an 81.5-second offset at the
    6.3$\sigma$ level.  The 1.3-day orbital period of WASP-4b appears
    to be decreasing at a rate of $\dot{P} = -12.6 \pm 1.2$
    milliseconds per year.  This might be caused by tidal orbital
    decay or apsidal precession, although both interpretations have
    shortcomings.  The gravitational influence of a third body is another
    possibility, although at present there is scant evidence for such a
    body. Further observations are needed to confirm and
    understand the timing variation.
\end{abstract}

@@

Although the Transiting Exoplanet Survey Satellite (TESS) is designed
to detect new planets, it will also perform precise photometric
monitoring of most of the planets that have been discovered by
ground-based transit surveys over the last two decades.  One of the
applications of the new TESS data is to search for timing anomalies in
hot Jupiter systems.  Long-term monitoring of hot Jupiter transit and
occultation times should eventually reveal two distinct processes.

First, whenever the orbital angular momentum is less than 3 times the
rotational angular momentum, the system is vulnerable to tidal orbital
decay \citep{counselman_outcomes_1973,hut_stability_1980}.  Most hot
Jupiters satisfy this condition and their orbits should be shrinking
\citep{levrard_falling_2009,matsumura_tidal_2010}.  The rate of decay,
which depends on how friction dissipates the energy of tidal
disturbances, is difficult to observe directly or calculate from first
principles. This has been a stubborn source of uncertainty in the
interpretation of many phenomena in stellar and planetary astrophysics
(as reviewed by \citealt{Mazeh2008} and \citealt{ogilvie_tidal_2014}).

Second, long-term timing studies may reveal apsidal precession, {\it
  i.e.}, the rotation of the orbital ellipse within the orbital
plane. Apsidal precession has long been observed in eclipsing binaries
\citep[{\it e.g.},][]{
  schwarzschild_structure_1958,torres_accurate_2010,borkovits_eclipse_2015}.
For exoplanets, precessing orbits have been detected in a handful of
cases, including circumbinary systems (Kepler-413
\citealt{kostov_kepler-413b_2014}; Kepler-453
\citealt{welsh_kepler_2015}, planets orbiting rapidly rotating stars
(Kepler-13
\citealt{szabo_spin-orbit_2012,szabo_mapping_2014},~\citealt{masuda_spin-orbit_2015};
and systems with multiple interacting giant planets (Kepler-108
\citealt{mills_kepler-108_2017}; Kepler-693
\citealt{masuda_eccentric_2017}).  Apsidal precession is not regularly
observed in hot Jupiter systems because their orbits are nearly
circular, which may itself be a consequence of star-planet tides.  If
it could be observed, then \citep{ragozzine_probing_2009} showed that
it would be possible to measure the planet's Love number, which would
be a valuable constraint on the planet's interior.

So far, the most convincing direct evidence for either orbital decay
or apsidal precession has come from WASP-12b.
\citet{maciejewski_departure_2016} and \citet{patra_2017} showed that
the transit period of WASP-12b has been steadily decreasing by about
30~milliseconds per year over the last decade.  Here, we present
evidence for a timing anomaly in the WASP-4 system.  The hot Jupiter
WASP-4b orbits a G7V star every 1.34 days, corresponding to an orbital
distance of 5.5 stellar radii
\citep{wilson_wasp-4b_2008,hoyer_tramos_2013}.  It is a good target to
search for departures from a constant period, because transit data
exists all the way back to 2007.  The orbital eccentricity is less
than 0.018 (2$\sigma$), based on the work of
\cite{knutson_friends_2014}, who combined the available transit times,
occultation times, and Doppler data.  The stellar obliquity (or at
least its sky projection) is also compatible with zero, within about
10 degrees
\citep{triaud_spin-orbit_2010,beerer_secondary_2011,sanchis-ojeda_starspots_2011}.

In what follows, \S~\ref{sec:observations} presents the new TESS
observations, and \S~\ref{sec:timing} describes our timing
analysis.  We tried fitting the data with three models: a constant
period, a decaying period, and a slightly eccentric precessing orbit.
A constant period can be ruled out, but we cannot distinguish between
the latter two possibilities.  Either possibility would have
interesting implications, described in \S~\ref{sec:implications}.
Appendix~\ref{sec:verify_tess} considers the possibility that the
WASP-4 timing anomaly is due to an error in time stamps in the TESS
data products.  We found this possibility to be unlikely because none
of the other hot Jupiters we examined show a timing offset with the
same amplitude as was seen for WASP-4.

\section{New transits and system parameters}
\label{sec:observations}

\begin{figure}[t!]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{f2.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
      {\bf TESS observations of WASP-4b.} On the left, black points
      are TESS flux measurements, with a vertical offset applied. Blue
      curves are best-fit models. The numbers printed next to each
      lightcurve are the approximate transit times expressed in
      BJD minus 2{,}457{,}000.  The right side shows the
      residuals.
       \label{fig:lightcurves}
    }
\end{figure}

\begin{figure}[t!]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{f1.png}
    \end{center}
    \vspace{-0.5cm}
    \caption{
        {\bf Phase-folded lightcurve of WASP-4b.} Black points are
        TESS data. Yellow points are binned measurements.  The bottom
        panel shows the residuals.  The fit to the phase-folded
        transit (blue line) is used when measuring mid-transit times for
        the individual transits (Figure~\ref{fig:lightcurves}).
        \label{fig:phasefold}
    }
\end{figure}

\subsection{Observations}

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network. The
timestamps from the spacecraft clock were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

We began our analysis with the Presearch Data Conditioning (PDC)
lightcurve \citep{smith_kepler_apertures_2017,smith_kepler_PDC_2017}.
We processed the lightcurve as follows.  First, we removed all points
with non-zero quality flags.  This removed data contaminated by coarse
spacecraft pointing, cosmic rays, and other annoyances
\citep{tess_data_product_description_2018}.  We also removed the data
within the first and last hours of both orbits, because of ramp-like
systematic effects that appear during those time ranges. We also
removed the data that might have been adversely affected by ``momentum
dumps,'' the firing of thrusters and resetting of reaction wheels that
took place every 2.5 days during sector
2.\footnote{\url{archive.stsci.edu/hlsps/tess-data-alerts/hlsp_tess-data-alerts_tess_phot_s01_tess_v1_release-notes.pdf}}.
The data during these events were assigned quality flags corresponding
to ``Reaction Wheel Desaturation Event'' and ``Manual Exclude''.  For
WASP-4, these flags were simultaneously set for 54 distinct cadences,
and there were 10 momentum dumps, averaging about 10 minutes of
flagged data per dump.  Out of caution, we clipped out an additional
10 minutes before and after every momentum dump.

All told, we removed 8\% of the original data points, and were left
with 18{,}165 measurements of the relative flux of WASP-4.  We
normalized the data by dividing out the median flux.  We converted the
timestamps from BTJD into BJD by adding the appropriate 2457000 day
offset \citep{tess_data_product_description_2018}.  Many of these and
subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}. We did not
``flatten'' the lightcurves, as is often done with splines,
polynomials, or Gaussian processes.  Instead, we modeled the
out-of-transit flux variations simultaneously with the transit parameters,
as described below.

@@

\begin{figure}[t!]
    \begin{center}
        \leavevmode
        \includegraphics[width=0.49\textwidth]{f3.pdf}
    \end{center}
    \vspace{-0.6cm}
    \caption{ {\bf TESS saw WASP-4b transit earlier than expected.}
      Both plots show the deviations between the observed and
      calculated transit times, where the calculation is based only on
      the pre-TESS data and assumes a constant period.  The blue bands
      depict the $\pm$$1\sigma$ credible interval of the predicted
      times.  {\it Top:} The full timing dataset spans 10 years. The
      darkest points correspond to the most precise data.  {\it
        Bottom:} Close-up of the TESS observations. The red band shows
      the average deviation of the TESS data, which is $81.5 \pm
      11.8\ {\rm seconds}$.
        \label{fig:arrived_early}
    }
\end{figure}

Using the cleaned PDC lightcurve, we used the Box Least Squares
algorithm \citep{kovacs_box-fitting_2002} to estimate the orbital
period, transit duration, and a reference epoch using the TESS data
alone.  Based on the results, we isolated the data within 4 transit
durations of each transit midpoint.  We then fitted all of the transit
data using the analytic formulae given by \citet{mandel_analytic_2002}
and implemented by \citet{kreidberg_batman_2015}.  We assumed the
orbit to be circular. The free parameters were the reference epoch,
the planet to star radius ratio $R_{\rm p}/R_\star$, the orbital
distance to stellar radius ratio $a/R_\star$, the inclination $i$, two
quadratic limb-darkening coefficients $(u_{\rm linear}, u_{\rm
  quad})$, and the orbital period $P$.
% JNW: Luke, didn't you also have to fit a linear trend for each transit?
% Otherwise the fit would be poor, for each transit, even if the
% trends tend to cancel out in the phase folded light curve.
% And, did you assign the photometric uncertainties as described below?

We sampled the posterior probability distribution for all the
parameters using the algorithm proposed by
\citet{goodman_ensemble_2010} and implemented by
\citet{foreman-mackey_emcee_2013}.  Table~1 gives the results, which
are in reasonable agreement with the parameters reported by {\it
  e.g.}, \citet{southworth_high-precision_2009} and
\citet{huitson_gemini_2017}.  Figure~\ref{fig:phasefold} shows the
phase-folded lightcurve.

To measure the transit times, we fitted each transit separately, with
four free parameters: the time of mid-transit $t_{\rm tra}$, the
planet-to-star radius ratio, and the slope and intercept of a linear
trend to account for any slow variations unrelated to the transit.  We
fixed the remaining parameters at the values that had been determined
from the phase-folded TESS lightcurve.  The uncertainty in each
data point was set equal to the root-mean-square (rms) level of the
out-of-transit data.

To verify that the measured uncertainties are plausible, we computed
the reduced $\chi^2$ for a linear ephemeris fit to the measured TESS
transit times.  We found that $\chi^2 = 9.2$, with $n=16$ degrees of
freedom.  The variance of the $\chi^2$ distribution is $2n$, so we
would expect $\chi^2 16 \pm 5.7$.  Visually inspecting the residuals
showed that the error variance had been overestimated, so we
multiplied the measured TESS errors by a factor $f=0.76$, forcing a
reduced $\chi^2$ of unity.  This lowered the mean uncertainty of the
transit midtimes from $29.8$ to $22.6$ seconds.

Figure~\ref{fig:lightcurves} shows the lightcurve of each transit, the
best-fit models, and the residuals.  Table~2 reports the mid-transit
times and their uncertainties.  After binning the residuals to 1-hour
windows, the lightcurves have an rms scatter of 586\,{\rm ppm}.  The
pre-launch TESS noise model \citep{winn_photonflux_2013,Sullivan_2015}
\footnote{\url{github.com/lgbouma/tnm}} would have predicted an error
budget consisting of the following terms added in quadrature:
410\,{\rm ppm} from photon-counting noise, 202\,{\rm ppm} from
detector read noise, and 673\,{\rm ppm} from the zodiacal background
light.  The level of background light appears to have been
overestimated.

@@

\tablecomments{
  (A) From phase-folded TESS lightcurve (\S~\ref{sec:measurement}).
  Orbital periods are in Table~4. The limb darkening parameters were
  allowed to float around the \citet{claret_limb_2017} prediction, but
  were unconstrained.
  (B) \citet{triaud_spin-orbit_2010}.
  (C) From HARPS spectra \citep{doyle_accurate_2013}.
  (D) \citet{stassun_accurate_2017}.
  (E) From functions of stellar, transit or RV parameters as discussed in text.
  (F) \citet{gaia_collaboration_gaia_2018}.
  (G) \citet{stassun_TIC_2018}.
}

@@

We calculated the stellar and planetary parameters in the following
way.  We computed the star's spectral energy distribution based on the
{\it Gaia} DR2 parallax (after making the small correction advocated
by \citealt{stassun_evidence_2018}) and the broadband magnitudes from
all the available all-sky catalogs: $G$ from {\it Gaia\/} DR2, $B_T$
and $V_T$ from {\it Tycho-2}, $BVgri$ from {\it APASS}, $JHK_S$ from
{\it 2MASS}, and the {\it WISE}~1--4 passbands, thus spanning the
wavelength range 0.4--22~$\mu$m.  We adopted the effective temperature
from the work by \citet{doyle_accurate_2013}, who determined the
spectroscopic parameters of WASP-4 using high-signal-to-noise
observations with the High Accuracy Radial-velocity Planet Searcher
(HARPS).  Then, we determined the stellar radius through the
combination of the bolometric luminosity and the effective
temperature, using the Stefan-Boltzmann law.  To determine the stellar
mass, we first computed the mean stellar density based on the value of
$a/R_\star$ that gave the best fit to the phase-folded TESS lightcurve
(for the relevant equation, see \citealt{seager_unique_2003} or
\citealt{winn_exoplanet_2010}).  The mass was calculated from the
radius and density, and the orbital distance was also calculated from
the radius and $a/R_\star$.  The planetary radius was calculated as
the product of $R_\star$ and $R_{\rm p}/R_\star$.  Finally, the planet
mass was calculated based on the stellar mass, the radial-velocity
amplitude observed by \citet{triaud_spin-orbit_2010}, and the orbital
inclination.

Table~1 gives the resulting parameters, which we adopted for the
remaining analysis.  Our parameters are in agreement with those of
previous investigators
\citep{wilson_wasp-4b_2008,gillon_discovery_2009,winn_transit_2009,southworth_homogeneous_2011,petrucci_no_2013}.
By comparing the star's luminosity and spectroscopic parameters with
the outputs of the Yonsei-Yale stellar-evolutionary models, we found
that WASP-4 is a main-sequence star with an age of approximately
$7\,{\rm Gyr}$.

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
		The residuals are the observed times minus the calculated times
		assuming a constant period.  The darkest points correspond to
		the most precise data. The constant-period model (gray line) is
		a poor description of the data.  Models with a uniformly
		decreasing period (blue) or apsidal precession (orange) provide
		better fits to the data.  The red square represents the
		combination of all the TESS data and is for display purposes
		only.  The models were fitted to all of the individual transit
		times.
		\label{fig:times}
	}
\end{figure*}

@@

Table~2 gives the transit times we used in our analysis. We included
data from peer-reviewed literature for which the analysis was based on
observations of a single, complete transit, and for which the midpoint
was allowed to be a free parameter. We also required that the time
system be clearly documented. Many of the times were previously
compiled by \citet{hoyer_tramos_2013}. We confirmed that the times in
that paper were in agreement with the original sources and that
barycentric corrections had been performed when needed.
